Safe Transferable Regions

### 2015-11-18 paper/overview2.tex
Edits to Sec.3.
Sec 3 full pass.
Edits to Sec.4. Formatting changes.

\subsection{Using Regions in \name}

@@

In \name, this enforcement is done at runtime by explicitly keeping
track of the \emph{current state} for $\RgnZ$ objects, and checking
the validity of every open, transfer, or free operation and throwing
an exception if it is invalid.  
% The challenge in enforcing this
% discipline statically is that transferable regions are first-class
% objects. Hence, the program can create multiple aliases for the same
% region, \eg, open it via one alias and free it via another.  Typestate
% verification in the presence of aliases is hard.  The checking can be
% done statically by preventing the creation of aliases using, \eg,
% linear types or unique types. However, this would be quite
% restrictive, in terms of expressiveness.
As explained previously, this is a reasonable tradeoff in the context
of \name, as regions are coarse-grained objects, which are manipulated
infrequently, when compared to fine-grained objects that reside inside
these regions. Therefore, runtime overhead of checking the region's
state transition discipline is acceptable.

\paragraph{Cloning} Note that in the example from
Fig.~\ref{fig:motivating-eg-in-broom} the object returned by the
\C{selector} (on Line 13) should not contain any references to the
input object, since the input region, where the object resides, will
be freed at the end of the method. If there is a need for the output
object to point to subobjects of the input object, such subobjects
must be cloned (to copy them from the input region to the output
region).  Fortunately, \name's region type system
(\S~\ref{sec:type-system}) is capable of capturing such nuances in the
type of \C{selector} and the type checker will ensure correctness.
Furthermore, the type can be automatically inferred by \name's region
type inference (\S~\ref{sec:type-inference}), which can peform the
above reasoning on behalf of the programmer.

>>>>

\subsection{Using Regions in \name}
\label{sec:alloc-ctxt}

@@

In \name, this enforcement is done at runtime by explicitly keeping
track of the \emph{current state} for $\RgnZ$ objects, and checking
the validity of every open, transfer, or free operation and throwing
an exception if it is invalid.  
% The challenge in enforcing this
% discipline statically is that transferable regions are first-class
% objects. Hence, the program can create multiple aliases for the same
% region, \eg, open it via one alias and free it via another.  Typestate
% verification in the presence of aliases is hard.  The checking can be
% done statically by preventing the creation of aliases using, \eg,
% linear types or unique types. However, this would be quite
% restrictive, in terms of expressiveness.
As explained previously, this is a reasonable trade-off in the context
of \name, as regions are coarse-grained objects, which are manipulated
infrequently, when compared to fine-grained objects that reside inside
these regions. Therefore, runtime overhead of checking the region's
state transition discipline is acceptable.

\input{fb-syntax}

\paragraph{Cloning} Note that in the example from
Fig.~\ref{fig:motivating-eg-in-broom} the object returned by the
\C{selector} (on Line 13) should not contain any references to the
input object, since the input region, where the object resides, will
be freed at the end of the method. If there is a need for the output
object to point to subobjects of the input object, such subobjects
must be cloned (to copy them from the input region to the output
region).  Fortunately, \name's region type system
(\S~\ref{sec:type-system}) is capable of capturing such nuances in the
type of \C{selector} and the type checker will ensure correctness.
Furthermore, the type can be automatically inferred by \name's region
type inference (\S~\ref{sec:type-inference}), which can perform the
above reasoning on behalf of the programmer.

### 2015-11-18 paper/related.tex
Progress lemma.
Edits to Sec.4. Formatting changes.

An ownership type system for safe region-based memory management in
real-time Java has been proposed by~\cite{MIT03}. Like us, they too
assume a source language with programmer-managed memory regions, and
focus on proving safety of programs written in that language. Their
source language admits various kinds of regions in order to support
shared-memory concurrency. However, all their regions have
lexically-scoped lifetimes. In contrast, we admit regions with
dynamically determined lifetimes in order to support message-passing
concurrency. We borrow outlives relation from their formal
development, and our type system bears some similarities to their's.
However, our language also admits parametric polymorphism (generics)
and higher-order functions, whose interaction is non-trivial in
context of a region type system. On the other hand, our type system
eschews the notions of region kinds and ownership, leading to a more
succinct formalization. Furthermore, we establish a type safety result
that formalizes its guarantees with respect to a well-defined
operational semantics. Like Cyclone, ~\cite{MIT03}'s language is
explicitly typed. Although there is some support for local type
inference, region types for methods and classes need to be written
explicitly. In contrast, we support full type inference that
eliminates any such need.

@@

The idea of using region-based memory management to facilitate the
safe transfer of rich data structures between computational nodes has
been previously explored by~\cite{gpu14} in the context of Scheme
language. However, their setting only includes lexically-scoped
regions for which Tofte and Talpin~\cite{tofte97} style analysis
suffices. In contrast, our language provides first-class support for
transferable regions with dynamic lifetimes. We require this
generality in order to support streaming query operators, such as the
one shown in Fig.~\ref{fig:motivating-eg}.

>>>>

An ownership type system for safe region-based memory management in
real-time Java has been proposed by~\cite{MIT03}. Like us, they too
assume a source language with programmer-managed memory regions, and
focus on proving safety of programs written in that language. Their
source language admits various kinds of regions in order to support
shared-memory concurrency. However, all their regions have
lexically-scoped lifetimes. In contrast, we admit regions with
dynamically determined lifetimes in order to support message-passing
concurrency. We borrow outlives relation from their formal
development, and our type system bears some similarities to theirs.
However, our language also admits parametric polymorphism (generics)
and higher-order functions, whose interaction is non-trivial in
context of a region type system. On the other hand, our type system
eschews the notions of region kinds and ownership, leading to a more
succinct formalization. Furthermore, we establish a type safety result
that formalizes its guarantees with respect to a well-defined
operational semantics. Like Cyclone, ~\cite{MIT03}'s language is
explicitly typed. Although there is some support for local type
inference, region types for methods and classes need to be written
explicitly. In contrast, we support full type inference that
eliminates any such need.

@@

The idea of using region-based memory management to facilitate the
safe transfer of rich data structures between computational nodes has
been previously explored by~\cite{gpu14} in the context of Scheme
language. However, their setting only includes lexically-scoped
regions for which Tofte and Talpin-style analysis~\cite{tofte97}
suffices. In contrast, our language provides first-class support for
transferable regions with dynamic lifetimes. We require this
generality in order to support streaming query operators, such as the
one shown in Fig.~\ref{fig:motivating-eg}.
